Rotation of Reynolds stress tensor [for shear-alignment]

### 2020-12-31 write/manuscript.tex
updated outer-layer profile for stream-wise shear-aligned component
ekman-profile for u,w in outer layer; modifications in inner w profile

@@

\end{itemize}
\begin{figure}
  \centerline{\includegraphics[width=0.7\textwidth]{ustar_alpha.pdf}}
  \caption{Variation of $Z_\star$ and surface veering with Reynolds number according to
    the theory by Spalart et al. (1989) and as estimated from DNS data}
  \label{fig:drag_law}
\end{figure} 
%
Attempts were also undertaken to obtain profiles of the wind speed: 
\cite{Gryning:2007dy} present an extension of the wind-speed profile beyond the surface layer using a neutral referece profile
and a stability correction;
\cite{Kelly:2010ev}, based on a probabilistic representation of stratification, develop a model for the long-term mean windspeed
in the atmospheric boundary layer and compare this with observation at different sites; 
\cite{Kelly:2016bs} demonstrate the effect of such improved model for wind-energy applications.
%
These investigations merely focus on the wind speed and thus do not regard the actual vector, i.e. the veering of
the wind with height is not described and there is little knowledge on the profile of the span-wise velocity component
and the precise shape of the hodograph in the limit of a truly neutral Ekman boundary layer.
%
While this is indeed an academic problem, the Ekman boundary layer often serves as limit for stratified boundary layers
and thus not only needs to be understood but also well-described as reference for higher-order approaches that take into
account possible effects of stratification, roughness or other physical complications of the geophysical system.
%
\subsection{Stream-wise Velocity} 
For the stream-wise velocity profile (that in non-rotating flows due to the geometry is always aligned with the surface shear stress),
well-established theories exist for various regimes according to their distance from the wall and the relative role of
viscosity, turbulence and interaction with the outer region of the flow with the logarithmic law for the mean velocity as a central anchor point:
\begin{subequations}
  \begin{itemize}
  \item In immediate vicinity to the surface, local turbulent mixing cannot occur for the no-slip/no-penetration boundary condition,
    and the mean velocity is described by a viscous profile of the form
    \begin{equation}
      (U,W)^{\alpha+}  = (z^+,0)  
    \end{equation}
    where the direction of 
    in absence of roughness elements are marginal linear regime / viscous sub-layer \cite{Foken:2002gv,Foken:1978bc} 
  \item buffer layer transition from linear regime to log-law  
  \item log-layer (K\'arm\'an, Prandtl, Jimenez, Marusic)
  \item outer log-layer () 
  \item wake region
  \item outer region 
  \end{itemize}
\end{subequations} 
%
\begin{figure}
  \includegraphics[width=\textwidth]{u_profile.pdf}
  \caption{Shear-aligned profiles of velocity components $U^{\alpha+}$ in inner (left) and outer (right) units.} 
\end{figure}

@@

%
\begin{figure}
  \includegraphics[width=1.0\textwidth]{w_profile.pdf}
  \caption{Profiles of shear-aligned span-wise velocity $(W^\alpha)^+$ versus inner and outer height.
  Dashed lines show DNS data, thick, opaque lines are from the semi-emprical theory developed above. }
  \label{fig:outer_w} 
\end{figure}
\end{subequations}

\begin{figure}
  \phantom{AAA}\textbf{(a)\hspace{0.3\textwidth}(b)\hspace{0.3\textwidth}(c)}\\
  \includegraphics[width=\textwidth]{ekman_profiles_data.pdf}
  \caption{ \textbf{(a)} Velocity deficity, \textbf{(b)} velocity profile in shear-aligned hodographs and \textbf{(c)} hodograph in geostrophy-aligned coordinates.
    Thick, solid lines show theory, dashed lines data from DNS. }
  \label{fig:hodograph} 
\end{figure}

@@

\textbf{Compare} to Emeis (2002) and Gryning (2007); highlight explicit knowledge on veering-profile $\rightarrow$ directional sheer;
%
\par%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
Implications for \textbf{K-theory} (we now can consider that shear and stress are not necessarily perfectly aligned).
$\rightarrow$ can we do something to infer a K-profile from these theoretical considerations? 
%
\begin{figure}
  \includegraphics[width=\textwidth]{mag_dir.pdf} 
  \caption{Total velocity and veering (in degrees) vs inner and outer height.
    Dashed lines show DNS data, thick lines are from semi-empirical theory.} 
\end{figure} 
%
\section{Conclusions}
%
Applications:
\begin{itemize}
\item reference-shear for neutral profile approaches(systematic!) $\rightarrow$ wind engineering! 
\item initial condition for LES/DNS to eliminate/minimize inertial oscillation in Benchmark simulations
\item 
\bibliographystyle{plainnat}
\bibliography{ansorge.bib}
\end{document}

>>>>

\newcommand{\p}{\partial}
\newcommand{\gzm}{{\xi}}%{\gamma z_\text{match}}}

@@

\end{itemize}
\begin{figure}
  \centerline{\includegraphics[width=0.7\textwidth]{../plot/ustar_alpha.pdf}}
  \caption{Variation of $Z_\star$ and surface veering with Reynolds number according to
    the theory by Spalart et al. (1989) and as estimated from DNS data}
  \label{fig:drag_law}
\end{figure} 
%
Attempts were also undertaken to obtain profiles of the wind speed: 
\cite{Gryning:2007dy} present an extension of the wind-speed profile beyond the surface layer using a neutral referece profile
and a stability correction;
\cite{Kelly:2010ev}, based on a probabilistic representation of stratification, develop a model for the long-term mean windspeed
in the atmospheric boundary layer and compare this with observation at different sites; 
\cite{Kelly:2016bs} demonstrate the effect of such improved model for wind-energy applications.
%
These investigations merely focus on the wind speed and thus do not regard the actual vector, i.e. the veering of
the wind with height is not described and there is little knowledge on the profile of the span-wise velocity component
and the precise shape of the hodograph in the limit of a truly neutral Ekman boundary layer.
%
While this is indeed an academic problem, the Ekman boundary layer often serves as limit for stratified boundary layers
and thus not only needs to be understood but also well-described as reference for higher-order approaches that take into
account possible effects of stratification, roughness or other physical complications of the geophysical system.
%
\subsection{Stream-wise Velocity} 
For the stream-wise velocity profile (that in non-rotating flows due to the geometry is always aligned with the surface shear stress),
well-established theories exist for various regimes according to their distance from the wall and the relative role of
viscosity, turbulence and interaction with the outer region of the flow with the logarithmic law for the mean velocity as a central anchor point:
\begin{subequations}
  \begin{itemize}
  \item In immediate vicinity to the surface, local turbulent mixing cannot occur for the no-slip/no-penetration boundary condition,
    and the mean velocity is described by a viscous profile of the form
    \begin{equation}
      (U,W)^{\alpha+}  = (z^+,0)  
    \end{equation}
    where the direction of 
    in absence of roughness elements are marginal linear regime / viscous sub-layer \cite{Foken:2002gv,Foken:1978bc} 
  \item buffer layer transition from linear regime to log-law  
  \item log-layer (K\'arm\'an, Prandtl, Jimenez, Marusic)
  \item outer log-layer () 
  \item wake region
  \item outer region 
  \end{itemize}
\end{subequations} 
%
\begin{figure}
  \includegraphics[width=\textwidth]{../plot/u_profile.pdf}
  \caption{Shear-aligned profiles of velocity components $U^{\alpha+}$ in inner (left) and outer (right) units.} 
\end{figure}

@@

%
\begin{figure}
  \includegraphics[width=1.0\textwidth]{../plot/w_profile.pdf}
  \caption{Profiles of shear-aligned span-wise velocity $(W^\alpha)^+$ versus inner and outer height.
  Dashed lines show DNS data, thick, opaque lines are from the semi-emprical theory developed above. }
  \label{fig:outer_w} 
\end{figure}
\end{subequations}

\begin{figure}
  \phantom{AAA}\textbf{(a)\hspace{0.3\textwidth}(b)\hspace{0.3\textwidth}(c)}\\
  \includegraphics[width=\textwidth]{../plot/ekman_profiles_data.pdf}
  \caption{ \textbf{(a)} Velocity deficity, \textbf{(b)} velocity profile in shear-aligned hodographs and \textbf{(c)} hodograph in geostrophy-aligned coordinates.
    Thick, solid lines show theory, dashed lines data from DNS. }
  \label{fig:hodograph} 
\end{figure}

@@

\textbf{Compare} to Emeis (2002) and Gryning (2007); highlight explicit knowledge on veering-profile $\rightarrow$ directional sheer;
%
\par%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
Implications for \textbf{K-theory} (we now can consider that shear and stress are not necessarily perfectly aligned).
$\rightarrow$ can we do something to infer a K-profile from these theoretical considerations? 
%
\begin{figure}
  \includegraphics[width=\textwidth]{../plot/mag_dir.pdf} 
  \caption{Total velocity and veering (in degrees) vs inner and outer height.
    Dashed lines show DNS data, thick lines are from semi-empirical theory.} 
\end{figure} 
%
\section{Conclusions}
%
Applications:
\begin{itemize}
\item reference-shear for neutral profile approaches(systematic!) $\rightarrow$ wind engineering! 
\item initial condition for LES/DNS to eliminate/minimize inertial oscillation in Benchmark simulations
\item
\end{itemize}

\appendix

\section{Laminar Ekman solution with consideration of inner layer}

\begin{subequations}
\begin{eqnarray}
  \left(\begin{matrix}
    \p_t U\\
    \p_t W  
  \end{matrix}\right)&=&\left(\begin{matrix}
     fW &+& \nu \p_z^2 U\\ 
    -f(U-G) &+& \nu \p_z^2 W
  \end{matrix}\right)\\ 
  \Rightarrow \partial_t (U+iW) &=& f(W-i(U-G))) + \nu \p_z^2(U+iW)
\end{eqnarray}
In stationary conditions, this system is solved by
\begin{eqnarray}
  \hat{u}(z) &=& U_{\inf} + e^{-\gamma z} \left[A \cos\gamma z + B \sin\gamma z\right]   \\
  \hat{w}(z) &=& W_{\inf} + e^{-\gamma z} \left[-A \sin\gamma z + B \cos\gamma z\right]
\end{eqnarray} 
\end{subequations}
where the constants $U_\inf$, $W_\inf$ set the top boundary condition and $A$ and $B$ set the bottom boundary condition. 
%
The most common boundary condition for a surface Ekman layer is $A=U_\inf=G$, $B=0$, and $W_\inf=0$.
%
The lower boundary condition, however, neglects the existence of the surface layer, and it appears reasonable to define
$A=c G$ where $c<1$ is a constant that icorporates the increased shear in the surface layer.
%
Given a 'matching height' $z_{match}$ and normalized matching height $\xi=\gamma z_\text{match}$ in the upper part of the inner layer, we can match the Ekman profile
to the inner layer by letting
%
\begin{eqnarray}
  \begin{matrix} 
  u(z_\text{match}) \equiv u_\text{match} &=& U_{\inf} + e^{-\gzm}} \left [ A \cos\gzm + B\sin\gzm \right] \\
    w(z_\text{match}) \equiv w_\text{match} &=& W_{\inf} + e^{-\gzm}} \left [-A \sin\gzm + B\cos\gzm \right]
  \end{matrix} 
\end{eqnarray}
\begin{eqnarray}
  e^\gzm  \left(\begin{matrix}
     u_\text{match}-U_{\inf} \\ 
     w_\text{match}-W_{\inf}
  \end{matrix}\right) = \left(\begin{matrix}
  \cos\gzm A &+ \sin\gzm B \\ -\sin\gzm A &+ \cos\gzm B
  \end{matrix} \right) \\
\end{eqnarray}
Matching the profile at $\xi=0$, one obtains
$A = \Delta u_\text{match} $ and $B=-\Delta w_\text{match} $. Otherwise, it is
\begin{subequations} 
\begin{eqnarray}
  e^\gzm \Delta u_\text{match} = A \cos\gzm + B\sin\gzm \\
  e^\gzm (\cos\gzm \Delta u_\text{match} - \sin\gzm \Delta w_\text{match}) = A (\cos^2\gzm + \sin^2\gzm)
\end{eqnarray}
\begin{eqnarray}
  \Rightarrow A &=&e^\gzm (\cos\gzm\Delta u_\text{match} + \sin\gzm \Delta w_\text{match})\\
  B&=&  \frac{e^\gzm \Delta u_\text{match} - A \cos \gzm}{\sin\gzm} \\
  &=& e^\gzm \frac{ \sin^2\gzm \Delta u_\text{match}+\sin\xi\cos\xi \Delta w_\text{match} }{\sin\xi }  \\
  &=& e^\gzm (\sin\xi\Delta u_\text{match} + \cos\xi\Delta w_\text{match}) 
\end{eqnarray}
\end{subequations} 
\bibliographystyle{plainnat}
\bibliography{ansorge.bib}

\end{document}
